---
title: "Demo Dependent t-test"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
  
Example 1: You have just been hired by a local non-profit that focuses on improving premature infant's weight, to see whether or not their parental education classes are effective. The weight of the premature infant is taken before parents learned strategies to help their premature infants and after the education classes. You collect a sample in your target population: low income families. Given this sample, what can you conclude about the parental education classes? Explain what this means to the head of the education program. Then explain this to someone who is well-versed in statistics. 

```{r}
library(psych)
library(jmv)
```

```{r}
dat_A <- read.csv("https://www.dropbox.com/s/155ouxhzbcjqw7f/dependent.A.csv?dl=1")
```

```{r}
desc <- descriptives(dat_A, vars = c('Before_Weight', 'After_Weight'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc
```


```{r}
dat_A$diff <- (dat_A$After_Weight - dat_A$Before_Weight)

desc <- descriptives(dat_A, vars = c('diff'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc
```


```{r}
ttestPS(data = dat_A, pairs = list(list(i1='Before_Weight', i2='After_Weight')), effectSize = TRUE, ci = TRUE, desc = TRUE)
```


Example 2: The non-profit wants to see if the education classes work in a different population. You collect the same size sample, but now with middle-upper class families. Given this sample, what can you conclude about the parental education classes? Explain what this means to the head of the education program. Then explain this to someone who is well-versed in statistics. 



```{r}
dat_B <- read.csv("https://www.dropbox.com/s/ht1asp6uqwf5z1w/dependent.B.csv?dl=1")
```


```{r}
desc.b <- descriptives(dat_B, vars = c('Weight_Before', 'Weight_After'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.b
```


```{r}
dat_B$diff <- (dat_B$Weight_After - dat_B$Weight_Before)

desc.b <- descriptives(dat_B, vars = c('diff'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.b
```


```{r}
ttestPS(data = dat_B, pairs = list(list(i1='Weight_Before', i2='Weight_After')), effectSize = TRUE, ci = TRUE, desc = TRUE)
```

Example 3: The nonprofit only has a little bit of money left for your studies, but they really want to see if their parental education classes are effective with just fathers. With the small amount of money, you can only collect a small sample, but you decide to run analyses anyway. Given this sample, what can you conclude about the parental education classes? Explain what this means to the head of the education program. Then explain this means to someone who is well-versed in statistics.

```{r}

dat_C <- read.csv("https://www.dropbox.com/s/awtdtf9cgwgf747/dependent.C.csv?dl=1")
```


```{r}
desc.c <- descriptives(dat_C, vars = c('Before_Weight', 'After_Weight'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.c

hist(dat_C$Before_Weight)
hist(dat_C$After_Weight)
```


```{r}
dat_C$diff <- (dat_C$After_Weight - dat_C$Before_Weight)

desc.c <- descriptives(dat_C, vars = c('diff'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.c

hist(dat_C$diff)
```


```{r}
ttestPS(data = dat_C, pairs = list(list(i1='Before_Weight', i2='After_Weight')), effectSize = TRUE, ci = TRUE, desc = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
