---
title: "PSY.308b.DA1"
output: word_document
---
  <!-- LEAVE THIS HERE, DO NOT PUT ANY CODE ABOVE IT -->
```{r, echo=FALSE, results=FALSE, message=FALSE, cache=FALSE}
library(knitr); opts_chunk$set(error=TRUE, cache=FALSE)

```

Instructions:
You are teaching your first Intro to Psychology course! After the first midterm you are disappointed with your students' overall test scores. You decide to implement two different required study techniques. There are 100 students in the class; 50 of them will be required to meet in groups to study right before the next midterm (Group A) and the other 50 will be required to create flashcards to aid in memorization (Group B). You are interested to know whether or not test scores improved significantly for each group, and you are also interested in which technique proved to be more beneficial. 

Hint: Use the PSYC308 Interpretation guide as a resource throughout your completion of the DA.


```{r}
#prep

library(psych)
library(jmv)

dat <- read.csv("https://www.dropbox.com/s/ouu2gswjn80leyt/PSY.308b.DA1.csv?dl=1")
```


```{r}
#Q2 

```

```{r}
#Q3 

```

```{r}
#Q4A 

#Q4B

```

```{r}
#Q7
 
```

Questions 

1.What are the two types of analyses you would need to use to answer your inquiries?


2. Please check the assumptions for the first type of analyses. Report whether each is violated or not and provide evidence to support your conclusions.


3. Please check the assumptions for the second type of analyses. Report whether each is violated or not and provide evidence to support your conclusions.

# dat.1 <- subset(dat_A, dat_A$Group == 'A')
# dim(dat_A)
4. You are interested in looking at the study techniques individually: 
(a) Did individuals in Group A obtain significantly higher exam scores after they engaged in the required study technique? 
(b) Did individuals in Group B obtain significantly higher exam scores after they engaged in the required study technique? 
Report only the relevant statistics according to APA format for both (a) and (b). Hint: don’t forget about the size of the effect…

5. The statistics professor in your psychology department heard about your little study you were conducting and was wondering what was going on with it. Interpret your findings from questions 4 (a) and (b) for the statistics professor. Hint: Use the PSYC308 Interpretation guide as a resource.


6. The Dean was also interested in your results as this may help to raise scores in other areas of the school too but unfortunately they do not understand statistical language. Please interpret your findings from questions 4 (a) and (b) for the Dean. Hint: Use the PSYC308 Interpretation guide as a resource.


7. Now test whether one of the study techniques was more effective than the other (Be sure to look at the test scores *after* they’ve engaged in their assigned study technique). Report all relevant statistics according to APA format. (Hint: don’t forget about the size of the effect…)


8. You call up Andrew since you know he would be so excited that you’re using statistics to help make decisions at work. Please explain what you found from question 7 to Andrew (who would also be stoked about the fun statistical jargon). Hint: Use the PSYC308 Interpretation guide as a resource.


9. Now that you have final findings from your analyses, please explain what you found to the Dean so that they can decide how to move forward with the other schools (remember - they do not understand statistical language well). Hint: Use the PSYC308 Interpretation guide as a resource.

10. In your own words, define homogeneity of variance. Why is it important? (Why should we consider it when doing certain analyses?).
