---
title: "<center>Mixed Factorial</center>"
author: "<center>MK</center>"
output:
  html_document:
    df_print: paged
---


```{r, echo=FALSE, results=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(knitr); opts_chunk$set(error=TRUE, cache=FALSE)
```

You're a researcher that is interested in the development of *self-control* in children.  You decide to examine the differences in the development of self-control among children who are in the presence of another child or alone. You randomly assign 30 children to two groups: either play in a room: a) with a peer or b) alone. You measure the length of time (in seconds) each child can wait before engaging with toys in the room. The measurement occurs at three different times: a) 1 year old (young), b) 3 (middle), and c) 5 (old) throughout their childhood.

*Independent Variables -*
*B-S - Room:* Peer, Alone
*R-M - Age:* Young, Middle, Old

*Dependent Variable -*
Time (Seconds)

```{r packages, include=FALSE}
library(psych)
library(reshape)
library(jmv)
library(car)
library(ggplot2)
```

```{r data}
dat <- read.csv("https://www.dropbox.com/s/a72gvqm4h23jamg/PSYC308B.MXF.Demo.csv?dl=1")

dat$ID <- as.factor(dat$ID) #Don't forget!

head(dat)

#order IV's correctly
levels(dat$Age)
dat$Age <- factor(dat$Age, levels = c("Young", "Middle", "Old"))
levels(dat$Age)
```

```{r}
# Assumptions: 

dat.wide <- cast(dat, ID + Room ~ Age, value = "Time")

# Normal distribution for R-M variable - 

dat.wide$diff1 <- (dat.wide$Young - dat.wide$Middle)
dat.wide$diff2 <- (dat.wide$Young - dat.wide$Old)
dat.wide$diff3 <- (dat.wide$Middle - dat.wide$Old)

descriptives(dat.wide, vars = c('diff1', 'diff2', 'diff3'), skew = TRUE, kurt = TRUE, hist = TRUE)

# Normal distribution for B-S variable -

# Need to find the average time for each B-S condition before being able to check the distribution with descriptives.

dat.wide$averagetime <- (dat.wide$Young + dat.wide$Middle + dat.wide$Old) / 3

descriptives(dat.wide, vars = c('averagetime'), splitBy = c('Room'), skew = TRUE, kurt = TRUE, hist = TRUE)

```

```{r Levenes Test}
# Levene's Test of Homogeneity of Variance

leveneTest(dat.wide$averagetime, dat.wide$Room, center = mean)
```

```{r}
#add bs and bsTerms 
model.rm <- anovaRM(data = dat.wide, 
                 rm = list(list(label = 'Age',
                                levels = c('Young', 'Middle', 'Old'))), 
                 rmCells = list(list(measure = 'Young', cell = 'Young'),
                                list(measure = 'Middle', cell = 'Middle'),
                                list(measure = 'Old', cell = 'Old')),
                 rmTerms = list('Age'),
                 bs = 'Room',
                 bsTerms = list('Room'),
                 effectSize = c('partEta'),
                 spherTests = TRUE,
                 spherCorr = c('none','GG'),
                 postHoc = list('Age'),
                 postHocCorr = 'holm',
                 emMeans = list(NULL))
model.rm

```

```{r Simple effects}
# Subset your data by Age
Young <- subset(dat, dat$Age == "Young")

Middle <- subset(dat, dat$Age == "Middle")

Old <- subset(dat, dat$Age == "Old")

#Example: we are running this for difference between the B-S conditions (room) at each level of R-M measure (young, middle, old).
ttestIS(data=Young, vars = 'Time', group = 'Room', eqv = T, effectSize = T)

ttestIS(data=Middle, vars = 'Time', group = 'Room', eqv = T, effectSize = T)

ttestIS(data=Old, vars = 'Time', group = 'Room', eqv = T, effectSize = T)

# What is another way that you could run simple effects? What would be the benefit of that?
# Try it on your own to find out!
```


```{r directionality}

# Young
descriptives(Young, vars = c('Time'), splitBy = c('Room'))

# Middle
descriptives(Middle, vars = c('Time'), splitBy = c('Room'))

# Old
descriptives(Old, vars = c('Time'), splitBy = c('Room'))
```


# Visualization (because we all love graphs)

```{r}
# again, we need to find the standard errors here so we can have beautiful error bars --- always have error bars...always.

summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

```{r}
sumdat <- summarySE(dat, measurevar="Time", groupvars=c("Room","Age"))

```

```{r}
# this plot with group them on the x by age and the cluster by room (three groups, with two bars)
plot <- ggplot(sumdat, aes(x = Age, y = Time, fill = Room))+
  geom_bar(stat='identity', position= 'dodge') +
  geom_errorbar(aes(ymin=Time-se, ymax=Time+se), position = position_dodge(.9), width = .1) +
  theme_minimal()

plot + ggtitle('Two-Way Interaction Graph for Self-Control by Room and Age')

```

```{r}
# ORRRRR you can have it this way where you look at each room on the x and they are clustered by age. Sometimes it's nice to see how the graphs may change based on how you want to present the data.

plot2 <- ggplot(sumdat, aes(x = Room, y = Time, fill = Age))+
  geom_bar(stat='identity', position= 'dodge') +
  geom_errorbar(aes(ymin=Time-se, ymax=Time+se), position = position_dodge(.9), width = .1) +
  theme_minimal()

plot2 + ggtitle('Two-Way Interaction Graph for Self-Control by Room and Age')

```