---
title: "One-way ANOVA Demo"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Example 1: 
A developmental psychologist is looking for a graduate student who can help them with their research project--you have been chosen! Their research looks into adolescent self-esteem and its relationship with parenting styles. There are three primary parenting styles: (1) authoritative (high expectations with support), (2) authoritarian (high expectations with little support), and (3) permissive (low expectations with support). You have collected data on 90 teenagers self-esteem levels and the parenting style they experienced (30 per style). Is there a significant difference in self-esteem by parenting style? What do your findings seem to indicate?

```{r}
#load library packages needed for analysis
library(psych)
library(jmv)
#load library packages needed for graphs
library(ggplot2)
library(plyr)
```

```{r}
#load data
dat_A <- read.csv("https://www.dropbox.com/s/vf50iy07fytkhjb/One-way%20ANOVA.A.csv?dl=1")

#View(dat_A) Row x Col
head(dat_A)
tail(dat_A)
dim(dat_A)

#encode $Style column as a factor because groups are numeric and need to be read as factor and not number
class(dat_A$Style)
dat_A$Style <- as.factor(dat_A$Style)
class(dat_A$Style)
```

```{r}
#Descriptives for all data, including histogram - does not check assumptions, but helps for big picture of data
desc <- descriptives(dat_A, vars = c('Self_esteem'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc

#Descriptives by group, including grouped histograms
desc.group <- descriptives(dat_A, vars = c('Self_esteem'), splitBy = 'Style', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.group

#Subset by group
dat1 <- subset(dat_A, dat_A$Style == "1")
dat2 <- subset(dat_A, dat_A$Style == "2")
dat3 <- subset(dat_A, dat_A$Style == "3")

#Histogram for each subsetted group
hist(dat1$Self_esteem)
hist(dat2$Self_esteem)
hist(dat3$Self_esteem)
```

```{r}
#anova with tukey post-hoc test
anova(data = dat_A, dep = 'Self_esteem', factors = c('Style'), effectSize = 'partEta', postHoc = c('Style'), postHocCorr = 'tukey', homo = TRUE)

```

```{r}
# we need to find the standard errors here so we can have beautiful error bars --- always have error bars...always.

summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

```{r}
sumdat <- summarySE(dat_A, measurevar="Self_esteem", groupvars=c("Style"))

```


```{r}
# creation of the bar graph - including specifications such as the color, title, addition of error bars, etc. 
bar1 <- ggplot(sumdat, aes(x = Style, y = Self_esteem)) +
  geom_bar(stat='identity', fill = 'lightblue2') +
  theme_minimal() +
  geom_errorbar(aes(ymin=Self_esteem-se, ymax=Self_esteem+se), width = .1)

bar1 + ggtitle('Self-Esteem by Parenting Style')

```

Example 2: You want to see if the trend is the same for younger children. You collected data on 60 kids' self-esteem levels and the parenting style they experienced (20 per style). Is there a significant difference in self-esteem by parenting style for these younger kids? What do your findings seem to indicate?


```{r}
dat_B <- read.csv("https://www.dropbox.com/s/fx9he863ywt5b56/One-way%20ANOVA.B.csv?dl=1")

dat_B$Style <- as.factor(dat_B$Style)
```

```{r}
desc.b <- descriptives(dat_B, vars = c('Self_esteem'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.b

desc.b.group <- descriptives(dat_B, vars = c('Self_esteem'), splitBy = 'Style', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.b.group

dat1.b <- subset(dat_B, dat_B$Style == "1")
dat2.b <- subset(dat_B, dat_B$Style == "2")
dat3.b <- subset(dat_B, dat_B$Style == "3")

hist(dat1.b$Self_esteem)
hist(dat2.b$Self_esteem)
hist(dat3.b$Self_esteem)
```

```{r}
anova(data = dat_B, dep = 'Self_esteem', factors = c('Style'), effectSize = 'partEta', postHoc = c('Style'), postHocCorr = 'tukey', homo = TRUE)

```

```{r}
# saving the summary data in an object
sumdat2 <- summarySE(dat_B, measurevar="Self_esteem", groupvars=c("Style"))

```


```{r}
#creation of the bar graph - including specifications such as the color, title, addition of error bars, etc. 
bar2 <- ggplot(sumdat2, aes(x = Style, y = Self_esteem)) +
  geom_bar(stat='identity', fill = 'orchid2') +
  theme_minimal() +
  geom_errorbar(aes(ymin=Self_esteem-se, ymax=Self_esteem+se), width = .1)

bar2 + ggtitle('Self-Esteem for Younger Children by Parenting Style')

```

Example 3:  You are interested in if the relationship seems to hold true even for teenagers that are a part of a mentoring program. There are only 30 teenagers involved mentoring program (10 per style). Is there a significant difference in their self-esteem by mentoring style? What do your findings seem to indicate?


```{r}

dat_C <- read.csv("https://www.dropbox.com/s/mq31b5zb8lh62qx/One-way%20ANOVA_C.csv?dl=1")


dat_C$Style <- as.factor(dat_C$Style)
```

```{r}
desc.c <- descriptives(dat_C, vars = c('Self_esteem'), hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.c

desc.c.group <- descriptives(dat_C, vars = c('Self_esteem'), splitBy = 'Style', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.c.group

dat1.c <- subset(dat_C, dat_C$Style == "1")
dat2.c <- subset(dat_C, dat_C$Style == "2")
dat3.c <- subset(dat_C, dat_C$Style == "3")


hist(dat1.c$Self_esteem)
hist(dat2.c$Self_esteem)
hist(dat3.c$Self_esteem)
```

```{r}


anova(data = dat_C, dep = 'Self_esteem', factors = c('Style'), effectSize = 'partEta', postHoc = c('Style'), postHocCorr = 'tukey', homo = TRUE)

```

```{r}
sumdat3 <- summarySE(dat_C, measurevar="Self_esteem", groupvars=c("Style"))

```


```{r}
# creation of the bar graph - including specifications such as the color, title, addition of error bars, etc. 
bar3 <- ggplot(sumdat3, aes(x = Style, y = Self_esteem)) +
  geom_bar(stat='identity', fill = 'lavender') +
  theme_minimal() +
  geom_errorbar(aes(ymin=Self_esteem-se, ymax=Self_esteem+se), width = .1)

bar3 + ggtitle('Self-Esteem for Teens by Parenting Style')

```
