---
title: "PSY.308b.FACTORIAL.DEMO"
output: html_document
---
 
  <!-- LEAVE THIS HERE, DO NOT PUT ANY CODE ABOVE IT -->
```{r, echo=FALSE, results=FALSE, message=FALSE, cache=FALSE}
library(knitr); opts_chunk$set(error=TRUE, cache=FALSE)

```

You are very interested in studying whether students's performance on essay-based exams differs depending on two different variables: (1) how many other students are in the room (i.e., alone, half, or a full class) and (2) whether they have breaks during the testing period (break; no break). You randomly select people from the college you attend to participate in your study (n = 120). You randomly assign people to one of the 6 conditions (alone-break, alone-no break, half-filled-break, etc.). Conduct the appropriate analyses to determine whether performance levels depend on these two variables with this sample.

DV: Score (1-10) - number of essay-formatted problems answered correctly.

#STUDY 1 
```{r}

library(psych)
library(jmv)
library(ggplot2)

dat_A <- read.csv("https://www.dropbox.com/s/7036my3n8lbb7nh/PSY.308b.FACTORIAL_DEMO_A.csv?dl=1")
```

#Assumptions
```{r}

dat_A.1 <- subset(dat_A, dat_A$Room == "Alone")
dat_A.2 <- subset(dat_A, dat_A$Room == "Full")
dat_A.3 <- subset(dat_A, dat_A$Room == "Half")

# Descriptives

desc.byroom <- descriptives(dat_A, vars = c('Score'), splitBy = 'Room', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.byroom

# OR you can try it this way

describeBy(dat_A, dat_A$Room)

hist(dat_A.1$Score)
hist(dat_A.2$Score)
hist(dat_A.3$Score)

```

```{r}

dat_A.4 <- subset(dat_A, dat_A$Break == "Break")
dat_A.5 <- subset(dat_A, dat_A$Break == "No Break")

desc.bybreak <- descriptives(dat_A, vars = c('Score'), splitBy = 'Break', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.bybreak

# OR this way...

describeBy(dat_A, dat_A$Break)

hist(dat_A.4$Score)
hist(dat_A.5$Score)
```

# ANOVA

```{r}
anova(data = dat_A, dep = 'Score', factors = c('Break', 'Room'), homo = TRUE)

#Levene's test
     #Residuals MS = within groups variance
     #dfTOTAL = #groups - 1 / (#groups (N-1)) = df1/df2

```

# ANOVA with post-hoc

```{r}


anova(data = dat_A, dep = 'Score', factors = c('Break', 'Room'), effectSize = 'partEta', postHoc = c('Break', 'Room'), postHocCorr = 'bonf')
```

# Simple effects

```{r}

#Simple effect of Break at each level of Room

# Alone
anova(data = dat_A.1, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')

# Half
anova(data = dat_A.2, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')

# Full
anova(data = dat_A.3, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')


```

#FYI

```{r}
#FYI


anova(data = dat_A.1, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')


ttestIS(dat_A.1, vars = 'Score', group = 'Break', eqv = TRUE, effectSize = TRUE, ci = TRUE)
```


```{r}


anova(data = dat_A, dep = 'Score', factors = c('Break', 'Room'), plotHAxis = 'Break', plotSepLines = 'Room')

```

# Visualization of Data using ggplot2 package.

```{r}
# Recommended that you save this function in a separate R script or doc to refer to later (trust me, it's worth it).

summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```

```{r}
sumdat <- summarySE(dat_A, measurevar="Score", groupvars=c("Break","Room"))

```

```{r}
plot1 <- ggplot(sumdat, aes(x = Room, y = Score, fill = Break))+
  geom_bar(stat='identity', position= 'dodge') +
  geom_errorbar(aes(ymin=Score-se, ymax=Score+se), position = position_dodge(.9), width = .1)

plot1 + ggtitle('Two-Way Interaction Graph of Performance Scores')

```

# But hey! Let's try something else...

You decide that those results were interesting, but you wonder what would happen if you randomly select from the top college in the US. You randomly select college students for your study (n = 110). You randomly assign people to one of the 6 conditions. Conduct the appropriate analyses to determine whether performance levels depend on these two variables with this sample.

#STUDY 2
```{r}
dat_B <- read.csv("https://www.dropbox.com/s/ke2vmbmokse96fj/PSY.308b.FACTORIAL_DEMO_B.csv?dl=1")
```

#ASSUMPTIONS
```{r}

dat_B.1 <- subset(dat_B, dat_B$Room == "Alone")
dat_B.2 <- subset(dat_B, dat_B$Room == "Half")
dat_B.3 <- subset(dat_B, dat_B$Room == "Full")

desc.byroomB <- descriptives(dat_B, vars = c('Score'), splitBy = 'Room', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.byroomB

# OR... (see the pattern here)...

describeBy(dat_B, dat_B$Room)

hist(dat_B.1$Score)
hist(dat_B.2$Score)
hist(dat_B.3$Score)
```

```{r}

dat_B.4 <- subset(dat_B, dat_B$Break == "No Break")
dat_B.5 <- subset(dat_B, dat_B$Braek == "Break")


desc.bybreakB <- descriptives(dat_B, vars = c('Score'), splitBy = 'Break', hist = TRUE, sd = TRUE, se = TRUE, skew = TRUE, kurt = TRUE)
desc.bybreakB

# OR...

describeBy(dat_B, dat_B$Break)

hist(dat_B.4$Score)
hist(dat_B.5$Score)

```


# ANOVA

```{r}
anova(data = dat_B, dep = 'Score', factors = c('Room', 'Break'), homo = TRUE)
```

# ANOVA with post-hoc

```{r}

anova(data = dat_B, dep = 'Score', factors = c('Room', 'Break'), effectSize = 'partEta', postHoc = c('Room', 'Break'), postHocCorr = 'bonf')
```

# Simple effects

```{r}


anova(data = dat_B.1, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')
anova(data = dat_B.2, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')
anova(data = dat_B.3, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')
```

#FYI

```{r}
#FYI

anova(data = dat_B.1, dep = 'Score', factors = c('Break'), effectSize = 'eta', postHoc = 'Break', postHocCorr = 'bonf')

ttestIS(dat_B.1, vars = 'Score', group = 'Break', eqv = TRUE, effectSize = TRUE, ci = TRUE)
```


```{r}

anova(data = dat_B, dep = 'Score', factors = c('Room', 'Break'), plotHAxis = 'Break', plotSepLines = 'Room')
```

# Visualization

```{r}
sumdat2 <- summarySE(dat_B, measurevar="Score", groupvars=c("Break","Room"))

```

```{r}
plot2 <- ggplot(sumdat2, aes(x = Room, y = Score, fill = Break))+
  geom_bar(stat='identity', position= 'dodge') +
  theme_minimal()+
  geom_errorbar(aes(ymin=Score-se, ymax=Score+se), position = position_dodge(.9), width = .1)

plot2 + ggtitle('Two-Way Interaction Graph of Performance Scores')

```

