---
title: "Network Modeling Tool"
author: "Christopher J. Shmank", "Daniel A. Pinedo"
date: "March 21, 2019"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #knitting markdown to html, word, or pdf
```

This document was adapted from a Workshop presented by Christopher J. Shmank
at Claremont Graduate University (CGU)

WARNING: the dataset we are using here uses a lot of processing power. Do not attempt this on a i386

TO-DO: 
  a. read qgraph library documentation
    - cor_auto()
    - qgraph()
    - averageLayout()
    - ggmFit()
  1. Find dataset with multiple correlations (e.g. Big Five in psych library)
  2. Clean dataset
  3. Apply qgraph function to dataset
  4. add useful comments
  5. add useful references

NOTE: **qgraph is the only loaded library currently implemented**

*Load libraries and data*
```{r}
library(pacman) #Package used to load all packages using p_load()
p_load (qgraph, NetworkComparisonTest, bootnet) 

#library(qgraph) #VISUALIZATION PACKAGE FOR GENERATING NETWORK MODELS
#library(NetworkComparisonTest) #USED TO DIRECTLY COMPARE TWO MODELS WITH SAME OBSERVED VARIABLES
#library(bootnet) #DETERMINE STABILITY OF CENTRALITY INDICES/NETWORK
```


```{r}
# Load big5 dataset:
data(big5)
data(big5groups)

FreqCOR<-cor_auto(big5) #create positive definite correlational matrix

# Correlations:
Q <- qgraph(FreqCOR, minimum = 0.25, cut = 0.4, vsize = 1.5, groups = big5groups, 
    legend = TRUE, borders = FALSE)
title("Big 5 correlations", line = 2.5)
```

```{r}

#Below is an example of groupings, here we are using the loaded "big5groups" from qgraph
#to group things in network graphs using pretty color groupings
#groups <- list( 
  #Example1=c(2, 5, 9, 13),
  #Example2=c(1, 4, 8, 12, 15),
  #Example3=c(3, 6, 11),
  #Example4=c(7, 10, 14) )

```

```{r}
g1 <- qgraph(FreqCOR, layout = "spring", groups = big5groups, title = "Big 5 Correlation Network", legend.cex = 0.5, details = TRUE)

pdf("g1.pdf", width=10, height=7) #output
qgraph(g1)
dev.off()
```

```{r}
g2 <- qgraph(FreqCOR, layout = "spring", graph = "pcor", groups = big5groups, title = "Big 5 Partial Correlation Network", legend.cex = 0.5, details = TRUE)

pdf("g2.pdf", width = 10, height = 7) #output
qgraph(g2)
dev.off()
```

```{r}
g3 <- qgraph(FreqCOR, layout = "spring", graph = "EBICglasso", sampleSize = 500, groups = big5groups, title = "Big 5 Regularized Correlation Network", legend.cex = 0.5, details = TRUE)

pdf("g3.pdf", width = 10, height = 7) #output
qgraph(g3)
dev.off()
```

```{r}
g3t <- qgraph(FreqCOR, layout = "spring", graph = "EBICglasso", threshold = TRUE, sampleSize = 500, groups = big5groups, title = "Big 5 Regularized Correlation Network (Threshold)", legend.cex = 0.5, details = TRUE)

pdf("g3t.pdf", width = 10, height = 7) #output
qgraph(g3t)
dev.off()
```

```{r}
gEBIC <- EBICglasso(FreqCOR, n = 9999, gamma = .5, lambda.min.ratio = .01)

pdf("gEBIC.pdf", width = 10, height = 7) #output
qgraph(gEBIC, layout = "spring", title = "EBICglasso Network", subtitle = "Gamma = 0.5, Lambda = 0.01", groups = groups)
dev.off()
```

```{r}
gBIC <- EBICglasso(FreqCOR, gamma = .0, lambda.min.ratio = .01, n = 9999)

pdf("gBIC.pdf", width = 10, height = 7) #output
qgraph(gBIC, layout = "spring", groups = groups, title = "BICglasso Network", subtitle = "Gamma = 0.0; Lamba = 0.01" )
dev.off()
```

```{r}
L <- averageLayout(g1,g2,g3)

pdf("Networks.pdf", width=10, height=7)
qgraph(g1, layout = L)
qgraph(g2, layout = L)
qgraph(g3, layout = L)

```

```{r}
# Fit Statistics of Network models
fitNetwork <- ggmFit(gEBIC, FreqCOR, sampleSize = 500, ebicTuning = .5)
fitNetwork2 <- ggmFit(gBIC, FreqCOR, sampleSize = 500, ebicTuning = .0)
fitNetwork3 <- ggmFit(g3, FreqCOR, sampleSize = 500, ebicTuning = .5)

fitNetwork
fitNetwork3
fitNetwork2
```

