---
title: "Moderation Analysis in Multiple Regression"
author: "Conway"
output:
  html_document: default
  pdf_document: default
---
  
  <!-- LEAVE THIS HERE, DO NOT PUT ANY CODE ABOVE IT -->
```{r, echo=FALSE, results=FALSE, message=FALSE, cache=FALSE}
library(knitr); opts_chunk$set(error=TRUE, cache=FALSE)
```

*Prompt*
----------
Suppose a clinical psychologist wanted to test the hypothesis that, among teenagers, state anxiety moderates the relationship between neuroticism and depression. The psychologist conducts a correlational study of 231 teenagers in a high school in the Midwest.

*Variables:*
bdi = Beck Depression Inventory
epiNeur = Neuroticism from Eysenck Personlaity Inventory
stateanx = state anxiety

*Data:*
These data are open source and available in R
From the console, View(epi.bfi)

```{r}

# Load packages
library(psych)
library(jmv)
library(car)
library(rockchalk)

# Create dataframe
dat <- epi.bfi

```

```{r}
# Descriptive statistics

desc <- descriptives(data = dat, 
                     vars = c('bdi', 'epiNeur', 'stateanx'), 
                     sd = TRUE, 
                     skew = TRUE, 
                     kurt = TRUE)
desc

# Correlations
round(cor(dat), 2)

# Correlations for this analysis
corvars <- c("bdi", "epiNeur", "stateanx")
newdat <- dat[corvars]
round(cor(newdat), 2)

```

```{r}
# Multiple regression, No moderation

model1 <- linReg(data = dat, 
                 dep = 'bdi', 
                 covs = c('epiNeur', 'stateanx'),
                 blocks = list('epiNeur', 'stateanx'), 
                 modelTest = TRUE, 
                 r2Adj = TRUE,
                 stdEst = TRUE,
                 ciStdEst = TRUE)
model1

```

```{r}
# Multiple regression, Moderation

dat$mod <- dat$epiNeur * dat$stateanx

model2 <- linReg(data = dat, 
                 dep = 'bdi', 
                 covs = c('epiNeur', 'stateanx', 'mod'),
                 blocks = list('epiNeur', 'stateanx', 'mod'), 
                 modelTest = TRUE, 
                 r2Adj = TRUE,
                 stdEst = TRUE,
                 ciStdEst = TRUE)
model2

```

```{r}
# Hierarchical regression wiht model comparison (R2 change when adding moderator)

modelcomp <- linReg(data = dat, 
                    dep = 'bdi', 
                    covs = c('epiNeur', 'stateanx', 'mod'),
                    blocks = list(
                      list('epiNeur', 'stateanx'), 
                      list('mod')), 
                    modelTest = TRUE,
                    r2Adj = TRUE,
                    stdEst = TRUE)
modelcomp

```

```{r}
# Moderation analysis with centered predictors

dat$epiNeur.c <- dat$epiNeur - mean(dat$epiNeur)
dat$stateanx.c <- dat$stateanx - mean(dat$stateanx)
dat$mod.c <- dat$epiNeur.c * dat$stateanx.c

modelcomp.c <- linReg(data = dat, 
                      dep = 'bdi', 
                      covs = c('epiNeur.c', 'stateanx.c', 'mod.c'),
                      blocks = list(
                        list('epiNeur.c', 'stateanx.c'), 
                        list('mod.c')), 
                      modelTest = TRUE, 
                      r2Adj = TRUE,
                      stdEst = TRUE,
                      ciStdEst = TRUE)
modelcomp.c

# Graph it
plot1 <- lm(bdi ~ epiNeur + stateanx + (epiNeur*stateanx), data = dat)
plotSlopes(plot1, plotx = "epiNeur", modx = "stateanx", modxVals = "std.dev.", main = "Moderation Analysis")

plot2 <- lm(bdi ~ epiNeur.c + stateanx.c + (epiNeur.c*stateanx.c), data = dat)
plotSlopes(plot2, plotx = "epiNeur.c", modx = "stateanx.c", modxVals = "std.dev.", main = "Moderation Analysis")


```

