---
title: "Binary Logistic Regression"
author: "Conway"
output:
  html_document: default
  pdf_document: default
---
  
  <!-- LEAVE THIS HERE, DO NOT PUT ANY CODE ABOVE IT -->
```{r, echo=FALSE, results=FALSE, message=FALSE, cache=FALSE}
library(knitr); opts_chunk$set(error=TRUE, cache=FALSE)
```

Prompt
----------
The data are based on a mock jury study conducted by Shari Diamond and Jonathan Casper. Subjects (N = 100) watched a videotaped sentencing phase trial in which the defendant had already been found guilty. The issue for the jurors to decide was whether the defendant deserved the death penalty. These data were collected "pre-deliberation"" (i.e., each juror was asked to provide his/her vote on the death penalty verdict, then the jurors met as a group to decide the overall jury verdict).The initial individual verdicts are given in this data set. Verdict is dummy coded: 0 = life sentence, 1 = death penalty.

**Load packages and import data**
```{r}

# Load packages
library(pacman)
p_load(psych, jmv, aod, QuantPsyc, ggeffects, ggplot2)

# Import data 
BL <- read.csv("https://www.dropbox.com/s/hd43va8a7hjfj27/Diamond.csv?dl=1")

```

**Descriptive statistics**
```{r}
# Descriptive statistics
# Including the binary outcome so we can see the frequencies (prints below Descriptives table)

desc <- descriptives(data = BL, 
                     vars = c('verdict', 'danger', 'rehab', 'punish', 'gendet', 'specdet', 'incap'), 
                     sd = TRUE, 
                     skew = TRUE, 
                     kurt = TRUE,
                     freq = TRUE,
                     hist = TRUE)
desc

```

**Correlations**
```{r}
# Correlation
cortable <- corrMatrix(data = BL, 
                       vars = c('verdict', 'danger', 'rehab', 'punish', 'gendet', 'specdet', 'incap'), 
                       flag = TRUE)
cortable
```

**Logistic Plots**
```{r}

p_load(popbio)
logi.hist.plot(BL$danger, BL$verdict, boxp=FALSE, type="hist", col="gray", xlabel = "danger")
logi.hist.plot(BL$rehab, BL$verdict, boxp=FALSE, type="hist", col="gray")
logi.hist.plot(BL$punish, BL$verdict, boxp=FALSE, type="hist", col="gray")
logi.hist.plot(BL$gendet, BL$verdict, boxp=FALSE, type="hist", col="gray")
logi.hist.plot(BL$specdet, BL$verdict, boxp=FALSE, type="hist", col="gray")
logi.hist.plot(BL$incap, BL$verdict, boxp=FALSE, type="hist", col="gray")

```

**BiLoRe Models**

```{r}

# Null model
model0 <- glm(BL$verdict ~ 1, family = binomial)
summary(model0)

model0.odds <- exp(coef(model0)) 
model0.odds
print("Odds")

model0.probs <- model0.odds / (1 + model0.odds)
print("Probabilities")
model0.probs

print("Columns = Observed, Rows = Predicted")
print("Null model")
ClassLog(model0, BL$verdict)
```

```{r}
# Model with predictors specific to the defendent.

model1.jmv <- logRegBin(
  data = BL,
  dep = verdict,
  covs = vars(danger, specdet, incap, rehab, punish, gendet),
  blocks = list(
    list(
      'danger',
      'specdet',
      'incap')),
  refLevels = list(
    list(
      var = 'verdict',
      ref = '0')),
  modelTest = TRUE,
  OR = TRUE,
  class = TRUE,
  acc = TRUE,
  collin = TRUE)

model1.jmv
```

```{r}
# Model with predictors about the criminal justice system. 

model2.jmv <- logRegBin(
  data = BL,
  dep = verdict,
  covs = vars(danger, specdet, incap, rehab, punish, gendet),
  blocks = list(
    list(
      'rehab',
      'punish',
      'gendet')),
  refLevels = list(
    list(
      var = 'verdict',
      ref = '0')),
  modelTest = TRUE,
  OR = TRUE,
  class = TRUE,
  acc = TRUE,
  collin = TRUE)

model2.jmv

```

```{r}
# Model with all predictors

model3.jmv <- logRegBin(
  data = BL,
  dep = verdict,
  covs = vars(danger, specdet, incap, rehab, punish, gendet),
  blocks = list(
    list(
      'danger',
      'specdet',
      'incap',
      'rehab',
      'punish',
      'gendet')),
  refLevels = list(
    list(
      var = 'verdict',
      ref = '0')),
  modelTest = TRUE,
  OR = TRUE,
  class = TRUE,
  acc = TRUE,
  collin = TRUE)

model3.jmv
```

**Model Comparison**

```{r}
# Model1 vs. Model3
## This is a similar set up to multiple regression in terms of the code, so if you know you will be running both models, you can just use this code and have the model comparison stats as well. 

modelcomp.jmv <- logRegBin(
  data = BL,
  dep = verdict,
  covs = vars(danger, specdet, incap, rehab, punish, gendet),
  blocks = list(
    list(
      'danger',
      'specdet',
      'incap'),
    list(
      'rehab',
      'punish',
      'gendet')),
  refLevels = list(
    list(
      var = 'verdict',
      ref = '0')),
  modelTest = TRUE,
  OR = TRUE,
  class = TRUE,
  acc = TRUE,
  collin = TRUE)

modelcomp.jmv
```

**Center Predictors**
```{r}

BL$dangerC <- BL$danger - mean(BL$danger)
BL$rehabC <- BL$rehab - mean(BL$rehab)
BL$punishC <- BL$punish - mean(BL$punish)
BL$gendetC <- BL$gendet - mean(BL$gendet)
BL$specdetC <- BL$specdet - mean(BL$specdet)
BL$incapC <- BL$incap - mean(BL$incap)

```

**Re-run model with all predictors and run parsimonious model.** 
```{r}

finalmodel.jmv <- logRegBin(
  data = BL,
  dep = verdict,
  covs = vars(dangerC, specdetC, incapC, rehabC, punishC, gendetC),
  blocks = list(
    list(
      'dangerC',
      'rehabC',
      'gendetC'),
    list(
      'specdetC',
      'incapC',
      'punishC')),
  refLevels = list(
    list(
      var = 'verdict',
      ref = '0')),
  modelTest = TRUE,
  OR = TRUE,
  class = TRUE,
  acc = TRUE,
  collin = TRUE)

finalmodel.jmv
```

# Use regression equation to calculate predicted logit, odds, and probability 
```{r}

# Let D = danger, R = rehab, G = gendet
D = 10
R = 0
G = 10

predlogit <- -.096 + (.278*D) + (-.181*R) + (.188*G)
predodds <- exp(predlogit)
predprob <- predodds / (1 + predodds)

predlogit
predodds
predprob

```
