---
title: "PSY.308d.DA4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

** IMPORTANT NOTES: PLEASE READ BEFORE STARTING **
The structure of your write-up should be: 
  - *Study 1 Results & Discussion:* typical results section + one paragraph discussion interpreting the results of your analyses
  - *Study 2 Results & Discussion:* typical results section + one paragraph discussion interpreting the results of your analyses
  - *Conclusion:* an INTEGRATED discussion of the two studies findings (do not simply repeat what has already been reported), implications, limitations, and future directions. 

**Additionally, all write-ups should be in APA format and your write-up should NOT exceed 4 pages.**

**Study 1: Pilot Study**
```{r}
dat <- read.csv("https://www.dropbox.com/s/qhx6mb45e1njvk4/PSY.308d.DA4_1.csv?dl=1")

#load libraries used
library(pacman)
p_load(jmv, psych, car, effsize)

#make dataset more readable
dat$GroupNum <- factor(dat$GroupNum,
  levels = c(1,2),
  labels = c("Not Hungry", "Very Hungry"))
```

McCormick & Co. is proposing a new chartiable foundation, FeedForward, which hopes to set-up meal program interventions in small towns to fight hunger amongst under-resourced citizens. While McCormick is a large company and rather successful, they want evidence before investing more money into the programs and other ideas. They recently hired you as an (sadly unpaid) intern to generate some data and give them the scoop and you have to do it on a budget (A.K.A. no funding for participants - just for supplies). You decide to put up some flyers in a couple nearby apartment buildings to recruit volunteers to join your pilot study for a meal program - luckily you got 9 people to participate!

*Research Question(s):*
1. Does the meal program significantly decrease the hunger compared to those on the waitlist? 
2. Does participants' weight significantly increase from baseline to the end of the program? 

*Variables:*
Score: self-reported hunger at the end of the program (1 = Not hungry, 100 = very hungry).
Group: program assignment (GroupNum; 1 = Program, 2 = Waitlist).
Pre: weight in pounds prior at baseline.
Post: weight in pounds at the end of two weeks.

Question #1
Assumptions:
1. Independence of observations
2. Normal distribution of dependent variable by condition
3. Homogeneity of variance

```{r}
# Descriptives by Condition
descriptives(data = dat,
             vars = 'Score',
             splitBy = 'GroupNum',
             sd = TRUE, 
             min = TRUE,
             max = TRUE,
             skew = TRUE,
             kurt = TRUE, 
             hist = TRUE)

hist(dat$Score [dat$GroupNum== 'Not Hungry'])
hist(dat$Score [dat$GroupNum== 'Very Hungry'])

```

```{r}

# Mann-Whitney test (with Levene's)
ttestIS(data = dat, 
        vars = 'Score',      # DV
        group = 'GroupNum',  # IV
        desc = TRUE,
        mann = TRUE,          # Mann-Whitney U
        eqv = TRUE,           # Levene's test
        meanDiff = FALSE,      
        ci = FALSE,
        effectSize = FALSE)   # Cohen's d

```

Effect Size
```{r}
# Cliff's Delta (non-parametric effect size) - a more robust version of Cohen's d which considers the ordinal nature, as opposed to interval, of most behavioral science and psych data (e.g., Likert scale). It is more powerful under conditions such as skewed distributions, etc. Essentially, it computes the dominance (overlap between two distributions) based on the probability that a selected score in Group 1 being greater than Group 2 minus the probability of a selected score in Group 1 being less than Group 2 divided by the product of the n of Group 1 and Group 2. The value range is between -1.0 and +1.0, where anything closer to the absolute value of 1 is no overlap (good) and closer to 0 is complete overlap (not so good).

cliff.delta(Score ~ GroupNum, data = dat, conf.level = .95, magnitude = TRUE, method = "Cliff's Delta")

# Because this package is nice - it gives you the Delta estimate and an interpretation.
```

Question #2
Assumptions:
1. Independence of observations
2. Normal distribution of dependent variable by time
3. Normal distribution of difference scores (Shapiro-Wilk)

```{r}
# Compute difference scores
dat$diff <- (dat$Post - dat$Pre)

# Descriptives
descriptives(data = dat, 
             vars = c('Pre', 'Post', 'diff'), 
             sd = TRUE, 
             range = TRUE, 
             min = TRUE, 
             max = TRUE, 
             skew = TRUE, 
             kurt = TRUE,
             hist = TRUE)

```

```{r}
 
# Wilcoxan Signed Rank Test (with Shapiro-Wilk)

#Pre/Post
ttestPS(dat, 
        pairs = list(
                list(i1 = 'Pre', i2 = 'Post')), 
        wilcoxon = TRUE, 
        norm = TRUE, 
        meanDiff = FALSE)

#Post/Pre
ttestPS(dat, 
        pairs = list(
                list(i1 = 'Post', i2 = 'Pre')), 
        wilcoxon = TRUE, 
        norm = FALSE, 
        meanDiff = FALSE)

```

Effect Size
```{r}
# Effect size for non-parametric related-samples t-test
# http://yatani.jp/teaching/doku.php?id=hcistats:wilcoxonsigned
# r = z/sqrt(N) where n is the total number of observations (cases x 2)

# To get number of cases(N):
n <- dim(dat)[1] # number of cases/rows
print(paste("Number of cases/rows(N):", N))

# To get sum of ranks(W)
W <- n*(n+1) / 2 # you can verify this by running Wilcoxon W in both directions and summing both scores
print(paste("Sum of ranks(W):", W))

# z = (W - mW + .5)/o
# https://www.statisticssolutions.com/how-to-conduct-the-wilcox-sign-test/
mW <- 0 #the null that there is no difference between T1 and T2

  #first we need to find sd  
  cat("\n")
  sd = sqrt((n*(n+1)*((2*n)+1))/6)
  print("sd = sqrt((n*(n+1)*((2*n)+1))/6)")
  print(paste("sd = sqrt((", n, "*(", n, "+1)*((2*", n, ")+1))/6)", sep = ""))
  print(paste("sd =", round(sd,2)))
 

  # now to find z:
  cat("\n")
  z = (W - mW + .5)/sd # technically this should be +- .5 dependening on directionality of W-score.
  print("z = (W - mW + .5)/sd")
  print("mW = the null that there is no difference between time 1 and time 2")
  print(paste( "z = (", round(W,2), " - ", mW, " + .5)/", round(sd,2), sep = ""))
  print(paste("z =", round(z,2)))

#r is the Pearson value reported for the non-parametric paired-sample t-test
cat("\n")
N = n*2 #total number of scores (2 x number of cases)
print(paste("Total number of scores(N):", N))
r = z/sqrt(N)
print("r = z/sqrt(n) = Effect size = the probability of a time 2 score being greater than a time 1 score")
print(paste("r = ", round(z,2), "/sqrt(", N, ")", sep = ""))
print(paste("r =", round(r,2)))

# Based on Cohen's criteria of evaluating effect size: .10 = small, .30 = medium, and .50 = large

# This z score above is iffy but was provided in full. R also calculates z scores for this test automagically, but I am not sure what is under the hood. See below

  # Wilcoxon signed rank test with z-score
  wilcoxon <- stats::wilcox.test(dat$Post, dat$Pre, paired = TRUE, exact = TRUE)
  wilcoxon.z <- stats::qnorm(wilcoxon$p.value) # z score estimate
  cat("\n")
  print("Wilcoxon test run from library(stats):")
  wilcoxon
  print(paste("Wilcoxon z-score run from library(stats):", round(wilcoxon.z, 2)))
  wilcoxon.r <- abs(wilcoxon.z/sqrt(N))
  print(paste("library(stats) effect size calculated as abs(", round(wilcoxon.z, 2), "/sqrt(", N, ")): ", round(wilcoxon.r, 2), sep = ""))
  
  # This runs what is called Exact Wilcoxon-Pratt Signed-Rank Test with a z-score
  library(coin)
  cat("\n")
  wilcoxon.coin <- coin::wilcoxsign_test(dat$Post ~ dat$Pre, distribution = "exact")
  wilcoxon.coin.z = as.numeric(statistic(wilcoxon.coin, type="standardized"))
  names(wilcoxon.coin.z) = "Z"
  print("Wilcoxon test run from library(coin):")
  wilcoxon.coin
  print(paste("Wilcoxon z-score run from library(coin):", round(wilcoxon.coin.z, 2)))
  wilcoxon.coin.r <- abs(wilcoxon.coin.z/sqrt(N))
  print(paste("library(coin) effect size calculated as abs(", round(wilcoxon.coin.z, 2), "/sqrt(", N, ")): ", round(wilcoxon.coin.r, 2), sep = ""))
  
```

**Study 2: Intervention**
```{r}
dat2 <- read.csv("https://www.dropbox.com/s/78onrh7icd92cs4/PSY.308d.DA4_2.csv?dl=1")
```

McCormick & Co. was impressed with your pilot study - meal programs seem to be worthwhile. They have invested $10,000 in your research account to do a bigger study where you can recruit and pay participants, gain more supplies, and extend your findings. You decide to set-up three different meal plan programs: Full Day (3 meals per day), Two-A-Day (2 meals per day), and One-A-Day (1 meal per day), and measure the average calorie intake above the noted minimum of 1,200 calories per day for each participant at the end of a two-week period. Given that these are under-resourced participants, they are encouraged to consume above the minimum calorie intake recommended by the health board. 

*Research Question(s):*
1. Is there a difference in the number of calories above the minimum consumed between the meal program conditions? 

