---
title: "Equity Valuation"
author: "Daniel Pinedo (from DataCamp course by Clifford Ang)"
date: "3/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Basics

After-Tax Income
[1] Revenues                         # Sales
[2] Less: Cost of Goods Sold
====================================
[3] Gross Profit
[4] Less: Operating Expenses
====================================
[5] Operating Income or EBIT         # Earnings Before Interest & Taxes
[6] Less: Interest Expense           # Compensation to debt holders
====================================
[7] Pre-Tax Income                   
[8] Less: Taxes                      # Payment to the government
====================================
[9] After-Tax Income                 # Net Income

FCFF model forecasts free cash flows for equity and debt and uses WACC
FCFE model forecasts free cash flows for equity accounting for debt and uses Cost of Capital

Time Value of Money
```{r}
# Present Value
     # Calculate Present Value of $100 n year from now given 5% interest rate of return
n = 1
r = .05

pv <- 100/(1 + r)^n
round(pv, 2)

```

FCFE model

"Free Cash Flows" are cash flows after you have paid out
     All your suppliers, employees, lenders, and government (taxes) and
     Setting aside money for capital investments and additional working capital needs and
     Net of new borrowings and debt repayments
     No effect on firm's projected operations
     No effect on firm's projected growth
     
Adjustments to Arrive to FCFE
[9] After-Tax Income
[10] Add: Depreciation and Amortization  # Non-cash charge. Cash spent at time of purchase
[11] Less: Capital Expenditures          # Cash spent on capital investments
[12] Less: Increases in Working Capital  # Cash spent on additional working capital needs
====================================
[13] Free Cash Flow to Equity

 






     


 
