---
title: "Mediation with Bootstrapping"
author: "Sir Stats McStatserson"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Mediation Demo with Bootstrapping**

A local college is hoping to understand what predicts how many events on average students will attend on campus per month. The board proproses that the best way to get higher attendance is to send a lot of e-mails. The more frequently that students are given information, the more they should attend. A student representative that sits in on board meetings believes there is more to the picture. While large weekly e-mail blasts may predict events attended on average, the interest a student has in attending the event should help explain how many events they attend. The board is looking to you to test these relationships. 

**Variables:**
Email - average e-mails read per week by students.
Interest - average amount of interest a student has for events listed in e-mails.
Events - average amount of events attended by students per month. 


Load in that data and those libraries.

```{r}

datboot <- read.csv('https://www.dropbox.com/s/qe8curup5safjy3/MedBootDemo.csv?dl=1')

library(pacman)
p_load(psych, jmv, medmod, lavaan, multilevel)

```

**Descriptives**

```{r}

desc <- descriptives(datboot, 
                     vars = c('Email', 'Interest', 'Events'), 
                     hist = TRUE, 
                     sd = TRUE, 
                     min = TRUE, 
                     max = TRUE, 
                     skew = TRUE, 
                     kurt = TRUE)
desc

```

**Correlation matrix**
```{r}

corr <- corrMatrix(datboot, 
                   vars = c('Email', 'Interest', 'Events'), 
                   flag = TRUE)
corr

```

**Regression**
```{r}

# X -> Y
model1 <- linReg(data = datboot, 
                 dep = 'Events', 
                 covs = 'Email',
                 blocks = list(c('Email')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE,
                 ciWidth = 95)
model1

# X -> M
model2 <- linReg(data = datboot, 
                 dep = 'Interest',
                 covs = 'Email',
                 blocks = list(c('Email')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE,
                 ciWidth = 95)
model2

# X + M -> Y
model3 <- linReg(data = datboot, 
                 dep = 'Events', 
                 covs = c('Email', 'Interest'),
                 blocks = list(c('Email', 'Interest')), 
                 modelTest = TRUE, 
                 stdEst = TRUE,
                 ci = TRUE, 
                 ciWidth = 95)
model3

```

**Mediation model - _without_ bootstrapping**
```{r}

# (a * b) = (c - c') = Indirect Effect [i.e., amount of mediation]
# Z  = Sobel test
# a  = Path Estimate from X to M
# b  = Path Estimate from M to Y
# c  = Total Estimate (Direct Estimate + Indirect Estimate)
# c' = Direct Estimate

med <- medmod::med(datboot, 
           dep = 'Events', 
           pred = 'Email', 
           med = 'Interest', 
           pm = TRUE, 
           paths = TRUE,
           label = TRUE,
           estPlot = TRUE)
med

```

**Mediation model - _with_ bootstrapping**
```{r}

boot <- medmod::med(datboot, 
            dep = 'Events', 
            pred = 'Email', 
            med = 'Interest', 
            estMethod = 'bootstrap', 
            bootstrap = 1000, 
            pm = TRUE, 
            paths = TRUE,
            label = TRUE,
            estPlot = TRUE)
boot

# Check the z-values, SE, and p-values for pathways to see differences (X -> Y for this one). 
```

### Sobel test

```{r}
# Look at this in comparison to the indirect mediation estimate.
# X = Email
# M = Interest
# Y = Events
sobel(datboot$Email, datboot$Interest, datboot$Events)

```




