---
title: "RegressionDemo.3"
author: "Conway"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Multiple Regression Demo - Centering, Dummy Coding, Suppressor Variable**

Prompt:

A large sample (N = 245) of middle-aged gym members were observed to examine the relationship between age, activity level, and endurance. Activity level was measured by the number of years the members had been physically active. Endurance was measured on a scale of 0 to 55 with higher values indicating greater endurance.

Variables:

age (X1)
activeyears (X2)
endurance (Y)

```{r echo = FALSE, message = FALSE}
library(psych)
library(car)
library(lsr)
library(jmv)
```


```{r}
dat <- read.csv("https://www.dropbox.com/s/asya9f3inm8zr1c/MR_EnduranceData.csv?dl=1")
```

**Descriptive Statistics and Assumptions**
```{r}
# Descriptives
desc <- descriptives(data = dat, 
                     vars = c('age', 'activeyears', 'endurance'), 
                     hist = TRUE, 
                     sd = TRUE, 
                     range = TRUE, 
                     skew = TRUE, 
                     kurt = TRUE)
desc
```

```{r}
# Scatterplots
plot(dat$age, dat$endurance, abline(lm(dat$endurance ~ dat$age)))
plot(dat$activeyears, dat$endurance, abline(lm(dat$endurance ~ dat$activeyears)))
plot(dat$age, dat$activeyears, abline(lm(dat$activeyears ~ dat$age)))
```

**Correlations**
```{r}
# Correlation
cortable <- corrMatrix(data = dat, 
                       vars = c('endurance', 'age', 'activeyears'), 
                       flag = TRUE)
cortable
```

**Simple Regression**
```{r}
# Simple regression
# R = correlation between observed scores and predicted scores
# R squared = percentage of variance explained
# t = Estimate / SE
# df = N - k - 1 [k is number of predictors]
# H0: B0 = 0; H0; R squared = 0
model1 <- linReg(data = dat, 
                 dep = 'endurance', 
                 covs = c('age'),
                 blocks = list('age'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE)
model1

model2 <- linReg(data = dat, 
                 dep = 'endurance',
                 covs = c('activeyears'),
                 blocks = list('activeyears'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE)
model2
```

**Multiple Regression**
```{r}
# Multiple regression
# Suppressor effect: active years is suppressing the relationship between age and endurance, and active years accounts for this
  #as evidenced by indication of standardized estimate going up when adding a predictor with age instead of remaining constant or going down as expected.
# In other words, for this sample, if you did not account for active years in relation to years, we are underestimating the prediction of age on endurance.
model3 <- linReg(data = dat, 
                 dep = 'endurance', 
                 covs = c('age', 'activeyears'),
                 blocks = list('age', 'activeyears'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE)
model3
```

**Model Comparison**
```{r}
# Model comparison
# H0 = delta of R squared = 0
compare <- linReg(data = dat, 
                  dep = 'endurance', 
                  covs = c('activeyears', 'age'),
                  blocks = list(
                    list('activeyears'),
                    list('age')), 
                  modelTest = TRUE, 
                  stdEst = TRUE, 
                  ci = TRUE)
compare
```

**Center the continuous predictor variables**
```{r}
# Centering only quantitatively changes the intercept for regression equation
# Center age and activeyears
dat$age.c <- dat$age - mean(dat$age)
dat$activeyears.c <- dat$activeyears - mean(dat$activeyears)
                                            
# Model comparison
compare2 <- linReg(data = dat, 
                  dep = 'endurance',
                  covs = c('activeyears.c', 'age.c'),
                  blocks = list(
                    list('activeyears.c'), 
                    list('age.c')), 
                  modelTest = TRUE, 
                  stdEst = TRUE, 
                  ci = TRUE)
compare2
```

**Invert age**
```{r}
#This is an unnecessary step.

# Invert age
dat$age.neg <- dat$age * -1

# Model comparison
compare <- linReg(data = dat, 
                  dep = 'endurance',
                  covs = c('activeyears', 'age.neg'),
                  blocks = list(
                    list('activeyears'), 
                    list('age.neg')), 
                  modelTest = TRUE,
                  stdEst = TRUE, 
                  ci = TRUE)
compare
```

**Let's add type of exercise to the dataframe**
```{r}
# Suppose there were three types of exercise: strength, aerobic, both
# Dummy code 1 (D1) will represent aerobic
# Dummy code 2 (D2) will represent strength
# The reference group will be both (D1 = 0, D2 = 0)
new.column <- vector(mode = "integer", length = 245)
dat$D1 <- new.column
for (i in 1:80)
{ dat$D1[i] <- 1}

new.column <- vector(mode = "integer", length = 245)
dat$D2 <- new.column
for (i in 81:160)
{ dat$D2[i] <- 1}

```



**Multiple regression with centered variables and dummy codes**
```{r}
# Model comparison
# D1 is predicted difference between D1 (aerobic) and reference group (both) for a 1 unit change in Y (endurance)
# D2 is predicted difference between D2 (strength) and reference group (both) for 1 unit change in Y (endurance)
compare <- linReg(data = dat, 
                  dep = 'endurance',
                  covs = c('activeyears.c', 'age.c', 'D1', 'D2'),
                  blocks = list(
                    list('activeyears.c', 'age.c'), 
                    list('D1', 'D2')), 
                  modelTest = TRUE,
                  stdEst = TRUE, ci = TRUE)

compare
```

