---
title: "MR_Demo2"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Multiple Regression Demo - Data Set #3**

Prompt:

A local psychologist has recruited you to help with a project investigating happiness in recent college graduates. The study collected a sample from the region of students who had graduated in the last 6 months based on a short survey. The psychologist was interested in what factors, including frequency of engagement with their community, number of hours worked, level of stress, optimism, and perceived social support influence happiness. The psychologist is interested these variables for a report coming out in the local newspaper about transitioning and adjustment after finishing college.

Variables:

Frequency - frequency of engagement with local community.
Hours - average number of hours worked per week.
Stress - average level of daily stress.
Opt - self-report level of optimism.
Support - rating of perceivied strength of social support.
Happiness - self-report level of happiness with life.

*Run the appropriate analyses and report relevant statistics in APA format. Please interpret the results for both someone who is knowledgable in statistics and for someone who is not.*

```{r echo = FALSE, message = FALSE}
library(psych)
library(car)
library(lsr)
library(jmv)
library(ggplot2)
library(ggeffects)
```

```{r}
dat <- read.csv("https://www.dropbox.com/s/gffxbm15lvaay2z/RegDemo.3.csv?dl=1")
```

**Descriptive Statistics and Assumptions**
```{r}
# Descriptives
desc <- descriptives(data = dat, 
                     vars = c('Frequency', 'Hours', 'Stress', 'Opt', 'Support', 'Happiness'), 
                     hist = TRUE, 
                     sd = TRUE, 
                     range = TRUE, 
                     skew = TRUE, 
                     kurt = TRUE)
desc
```

```{r}
# Scatterplots
plot(dat$Frequency, dat$Happiness, abline(lm(dat$Happiness ~ dat$Frequency)))
plot(dat$Hours, dat$Happiness, abline(lm(dat$Happiness ~ dat$Hours)))
plot(dat$Stress, dat$Happiness, abline(lm(dat$Happiness ~ dat$Stress)))
plot(dat$Opt, dat$Happiness, abline(lm(dat$Happiness ~ dat$Opt)))
plot(dat$Support, dat$Happiness, abline(lm(dat$Happiness ~ dat$Support)))
```

```{r}
# Homoscedasticity

ncvTest(lm(Happiness ~ Frequency + Hours + Stress + Opt + Support, data = dat))

```


**Correlations**
```{r}
# Correlation
cortable <- corrMatrix(data = dat, 
                       vars = c('Frequency', 'Hours', 'Stress', 'Opt', 'Support', 'Happiness'), 
                       flag = TRUE)
cortable
```

**Simple Regression**
```{r}
# Simple regression
model1 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Frequency'), 
                 blocks = list('Frequency'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE, 
                 ciWidth = 95)
model1

model2 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Hours'),
                 blocks = list('Hours'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE, 
                 ciWidth = 95)
model2

model3 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Stress'), 
                 blocks = list('Stress'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE, 
                 ciWidth = 95)
model3

model4 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Opt'),
                 blocks = list('Opt'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE, 
                 ciWidth = 95)
model4

model5 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Support'),
                 blocks = list('Support'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE, 
                 ciWidth = 95)
model5
```

**Multiple Regression**
```{r}
# Multiple Regression
model6 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Stress', 'Opt'),
                 blocks = list(c('Stress', 'Opt')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model6

model7 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Stress', 'Opt', 'Support'), 
                 blocks = list(c('Stress', 'Opt', 'Support')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model7

model8 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Stress', 'Opt', 'Support', 'Frequency', 'Hours'),
                 blocks = list(c('Stress', 'Opt', 'Support', 'Frequency', 'Hours')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model8
```

**Model Comparison**
```{r}
# Hierarchical regression with model comparison (R2 change)
model9 <- linReg(data = dat, 
                 dep = 'Happiness', 
                 covs = c('Stress', 'Opt', 'Support'),
                 blocks = list(
                   list('Stress', 'Opt'), 
                   list('Support')), 
                 modelTest = TRUE, 
                 r2Adj = TRUE, 
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model9
```

**Interpretation**
Always remember to interpret for a stats-knowledgable and not-very-statsy person!

**Visualization** 
```{r}
# Example of plotting a multiple regression model based on stress, optimism, and support predicting happiness
# save regression equation in an object
fit1 = lm(Happiness ~ Stress + Opt + Support, data = dat)
summary(fit1)

# create predicted values from three predictors and save in object
model_p <- ggpredict(fit1, terms = c('Stress', 'Opt', 'Support'), full.data = TRUE,  pretty = FALSE)

# plot predicted line
plot <- ggplot(model_p, aes(x, predicted)) +
      geom_smooth(method = "lm", se = FALSE, fullrange=TRUE) + xlab("Score") + ggtitle("Plot of Model Predicting Happiness") + ylab("Predicted Happiness") +
      geom_point() + theme_minimal()

plot 
```

**Dummy-Coded Variables Example!**

*Prompt:*

The same university was interested in assessing the number of hours faculty members worked in a given week and how that may be predicted by a variety of indicators. The HR department collected some preliminary data such as employment level (adjunct, tenure-track), salary, sex (female, male), perceived level of departmental support, and age. The department is specifically interested in salary, perceived departmental support, and employment level. The psychology department suggested you, an amazing statistics student, assess the data and report back to them in the next week. 

*Variables:*

Hours - average number of hours worked per week (numeric; continuous).
Salary - salary reported for the most recent academic year (in thousands).
Level - employment level in the departmet (numeric - dummy-coded; adjunct = 0, tenure-track = 1).
Age - in years (numeric; continuous).
Sex - (numeric - dummy-coded; female = 0, male = 1).
Support - perceived level of departmental support (numeric; continuous: 1-10 scale, where higher scores indicate greater perceived support).

*Run the appropriate analyses and report relevant statistics in APA format. Please interpret the results for both someone who is knowledgable in statistics and for someone who is not.*

```{r}
dat2 <- read.csv('https://www.dropbox.com/s/8txfj07k0obpfs3/MR_DummyData.csv?dl=1')

```


**Descriptive Statistics and Assumptions**
```{r}
# Descriptives
desc2 <- descriptives(data = dat2, 
                     vars = c('Hours', 'Salary', 'Level', 'Support'), 
                     hist = TRUE, 
                     sd = TRUE, 
                     range = TRUE, 
                     skew = TRUE, 
                     kurt = TRUE)
desc2
```

```{r}
# Scatterplots
plot(dat2$Salary, dat2$Hours, abline(lm(dat2$Hours ~ dat2$Salary)))
plot(dat2$Support, dat2$Hours, abline(lm(dat2$Hours ~ dat2$Support)))

```

```{r}
# Homoscedasticity

ncvTest(lm(Hours ~ Salary + Level + Support, data = dat2))

```

**Correlations**
```{r}
# Correlation
cortable2 <- corrMatrix(data = dat2, 
                       vars = c('Hours', 'Salary', 'Support', 'Age', 'Sex', 'Level'), 
                       flag = TRUE)
cortable2
```

**Regression**
```{r}
# Simple regression
model1.b <- linReg(data = dat2, 
                 dep = 'Hours', 
                 covs = c('Salary'), 
                 blocks = list('Salary'), 
                 modelTest = TRUE, 
                 stdEst = TRUE, 
                 ci = TRUE, 
                 ciWidth = 95)
model1.b
```

```{r}
# Multiple regression with comparison
model2.b <- linReg(data = dat2, 
                 dep = 'Hours', 
                 covs = c('Support', 'Salary'),
                 blocks = list(
                   list('Salary'),
                   list('Support')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model2.b 
```

```{r}
model3.b <- linReg(data = dat2, 
                 dep = 'Hours', 
                 covs = c('Support', 'Salary', 'Level'),
                 blocks = list(
                   list('Salary', 'Support'),
                   list('Level')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model3.b

```

**What about centering?**

```{r}
#Support
dat2$Support.C <- dat2$Support - mean(dat2$Support)
  
#Salary
dat2$Salary.C <- dat2$Salary - mean(dat2$Salary)

```

```{r}
# Multiple regression with comparison
model2.b.centered <- linReg(data = dat2, 
                 dep = 'Hours', 
                 covs = c('Support.C', 'Salary.C'),
                 blocks = list(
                   list('Salary.C'),
                   list('Support.C')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model2.b.centered
```

```{r}
model3.b.centered <- linReg(data = dat2, 
                 dep = 'Hours', 
                 covs = c('Support.C', 'Salary.C', 'Level'),
                 blocks = list(
                   list('Salary.C', 'Support.C'),
                   list('Level')), 
                 modelTest = TRUE,
                 r2Adj = TRUE,
                 stdEst = TRUE, 
                 ciStdEst = TRUE, 
                 ciWidth = 95)
model3.b.centered

``` 